Methods

### 2020-09-24 text/main.Rmd
commit before deep clean

# Trade-offs redefine the vulnerability of reef fish functions

Nina M. D. Schiettekatte, Simon J. Brandl, Jordan M. Casey, Nicholas A. J. Graham, Diego R. Barneche, Deron E. Burkepile, Jacob E. Allgeier, Katrina S. Munsterman, Fabien Morat, Alexandre Mercière, Yves Letourneur, Enrico L. Rezende, Graham J. Edgar, Rick D. Stuart-Smith, Jesús E. Arias-Gonzaléz, Alan M. Friedlander, Alison L. Green, Osmar J. Luiz,  F. A. Rodríguez‐Zaragoza, Laurent Vigliola, Michel Kulbicki, Sergio R. Floeter, Sébastien Villéger, Valeriano Parravicini

**Coral reefs flourish under nutrient-poor environments, and host a high diversity of fishes providing nutrition to millions of people [@Teh2013]. Understanding the ability of coral reefs to maintain healthy functioning is essential to conserving coral reefs and the services they provide to humanity, especially under the escalating pressures of fishing [@Graham2017; @Ceccarelli2016: @Stuart-Smith2018] and climate change [@Hughes2017]. Here, we characterize the predictors as well as the vulnerability of five key functions—nitrogen and phosphorus cycling, biomass production, herbivory, and piscivory—mediated by reef fishes across the world’s tropical oceans. We demonstrate that critical trade-offs exist among functions, such as between biomass production and phosphorous cycling, with no single reef fish community sustaining high values for all five functions simultaneously. In addition, we reveal that there is a high degree of functional dominance across communities. However, the identity of these species varies significantly across reefs, with ~70% of all species contributing disproportionally to a function in a given community. Finally, we highlight unbalanced vulnerability to human stressors across functions. Our results introduce a conundrum for coral reef conservation, i.e. we cannot maximize all functions, and imply that the successful preservation of coral reef functioning requires both conserving high fish diversity globally, and applying locally administered actions. **

The flow of elements through biological communities represent key ecosystem functions that fuel all life on Earth [@Welti2017]. There is a strong consensus on the importance of preserving these fluxes, also defined as ecosystem functions [@Welti2017]. 
However, understanding the dynamics of multiple ecosystem functions simultaneously is challenging because the drivers, vulnerability, and trade-offs among functions are still poorly characterized, impeding informed management [@Brandl2019front].
For millennia, humans have managed nature with an economic mindset to maximize desirable functions such as plant or animal biomass production [@Weisser2017]. 
However, trade-offs likely exist among ecosystem functions with efforts to maximize one function possibly negatively impacting another (e.g. timber vs. erosion control) [@Tallis2008]. 
To make informed management decisions that conserve ecosystem functioning, it is critical to quantify multiple functions simultaneously in order to understand the mechanisms that drive their potential trade-offs [@Rodriguez2006].

Coral reefs are among the most diverse and productive ecosystems on earth, and they provide indispensable ecosystem services to humans [@Teh2013]. 
As coral reefs thrive in nutrient-poor waters, efficient nutrient recycling is key to their productivity [@Brandl2019science]. 
 However, the integrity of coral reefs is threatened by a plethora of anthropogenic stressors, such as intensive exploitation and climate change [@Hughes2017]. 
Over the past decade, severe biodiversity declines have brought coral reef functioning and services to the forefront of scientific discourse [@Williams2019; @Bellwood2019; @Brandl2019front].
Yet, our capacity to quantitatively evaluate, monitor, and compare reef functioning primarily relies on static proxies of functions, such as fish standing biomass, and functional richness based on qualitative species traits [e.g. @Mora2011plos; @Cinner2020; @Mouillot2014]. 
Consequently, we know little about the co-occurrence of functions and their drivers, which currently hinders our ability to provide information that could be useful to coral reef management in a rapidly changing world [@Allgeier2016].

Here, we integrate, for the first time, biogeochemistry and community ecology to advance understanding and conservation of the fluxes of nutrients that underpin reef functioning. 
Specifically, we use a combination of novel data and phylogenetic extrapolation techniques to parametrize individual-level bioenergetic models which output five key ecosystem functions. 
We apply these models for 1,010 reef fish species and integrate their output functions with community structure data from 9,118 reefs at 585 sites across the world’s coral reefs to: 
(1) quantify nitrogen (N) excretion, phosphorus (P) excretion, biomass production, herbivory, and piscivory, 
(2) determine if there are trade-offs among these functions, 
(3) determine the ecological predictors of variation in ecosystem functions using Bayesian mixed models, 
(4) estimate the contribution and dominance of different species to each function, and 
(5) assess the vulnerability of these functions to climate change and fishing.

Combining novel approaches enabled us to quantify five key ecosystem functions performed by fishes across the world's reefs for the first time, highlighting the functional variation that exists beyond biomass effects. 
Biomass is the most commonly employed indicator of coral reef functioning [@MacNeil2015, @Mora2011plos; @Cinner2020; @Brandl2019front], and we observed a strong relationship between biomass and all five functions. 
Yet, correcting each function for biomass shows a high variability (Fig. 1), and for communities with similar total fish biomass, different functions often varied by over two orders of magnitude (Extended Data Fig. 1a). Further, we revealed strong trade-offs among the different biomass-corrected functions (Fig.1, Extended Data Fig. 1b). 
For example, communities with high herbivory rates or N excretion were negatively correlated with rates of P excretion.
Strikingly, less than 3% of all communities had an above average value of biomass-corrected function for all five functions simultaneously, and not a single community reached maximum values (upper 10%) for all functions. 
As a consequence, for a fixed amount of standing biomass, no reef appears to sustain high values for all of the five functions analyzed here, i.e. while many reefs may stand out as hotspots for one function, they may also output less of other functions (Fig. 1).

To uncover what underpins the trade-offs between multiple functions on coral reefs, we explored the relationship of each function with ecological predictors describing fish community structure and representing the variation of fish traits that affect elemental fluxes [@Schiettekatte2020] including body size, trophic level, species richness, and age structure (expressed as immaturity) by fitting Bayesian mixed models (Fig. 2). 
We found that negative correlations between functions are mediated by contrasting aspects of community structure (Fig. 2). 
For example, P excretion is higher in communities with a high proportion of large-bodied, mature fishes that occupy high trophic levels, thus creating positive relationships between P excretion and piscivory (Extended Fig. 1b) (See also ref. @Allgeier2014). 
In contrast, biomass production is highest in communities dominated by small, immature fishes at lower trophic levels, creating a strong trade-off between biomass production and P excretion. 
Metabolic theory predicts that small individuals have a higher mass-specific metabolic rate, and as a consequence, small individuals have elevated consumption rates and disproportionally contribute to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob].
On the other hand, fishes in early life stages that are highly dependent on phosphorus to grow, and fishes with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions to phosphorus excretion. 
These results vindicate the need to integrate size scaling and species-specific biological traits at the individual level with community ecology to understand ecosystem-wide elemental fluxes [@Schramski2015].

Human impacts have caused severe global declines of reef fish biomass and shifts in community structure [@Mora2011plos; @Hughes2017; @Williams2019], and our findings demonstrate that these changes will continue to fundamentally alter ecosystem functioning on reefs. 
For example, intensive fishing leads to reductions in biomass, which particularly selects for large fishes and consequently truncates the size, age, and trophic structure of fish communities [@Graham2017]. 
These effects can favor biomass-corrected N excretion and production [e.g. @Morais2020glob], while negatively impacting phosphorus excretion, herbivory, and piscivory (Fig. 2, Extended Data Fig. 2). 
On the other hand, changes in community structure triggered by coral cover declines related to climate change are associated with a shift toward herbivores, which may help avoid algal domination [@Graham2006; @Ceccarelli2016]. 
In contrast, herbivores have a minor contribution to phosphorus excretion [@Allgeier2014; @Schiettekatte2020], so a shift to herbivore dominance in fish communities and the subsequent decline of community-level phosphorus excretion may change the balance of nutrient cycling on reefs. 
Higher N:P ratios can in fact favor algal growth [@Burkepile2013], and promote symbiont dominance within the coral holobiont [@Allgeier2020].
This is particularly problematic when considered in conjunction with climate change, because phosphorus starvation in corals reduces their temperature threshold for bleaching [@Ezzat2016]. 
Considering multiple functions thus paints a more nuanced and complete picture to help us understand how human-induced shifts in community structure affect coral reef health.

Alongside features of community structure, the vulnerability of specific species that play a dominant role affects the vulnerability of community-level functions to human stressors such as climate change and fishing. 
Due to species-specific characteristics, certain species or entire families tend to contribute more or less to a particular function, relative to their contribution to standing biomass (Fig. 3a).
Further, abundant species can profoundly affect rates of functioning on the community level [@Ruttenberg2019; @Topor2019]. 
Therefore, we assessed the role of species by quantified the degree of dominance per function for each community. The degree of dominance describes the extent to which a function is performed by a single or only a couple of species. 
We demonstrate that for most communities, functions are consistently dominated by a few species (Fig. 3b).
We then asked if species consistently perform a key role when present across communities by quantifying the frequency of their dominance.
Despite evidence for species assuming key functional roles across widely-dispersed locations [@Bellwood2003], the identity of functionally-dominant species varied remarkably and the frequency of dominance per species tends to be low (Fig. 3c). 
Consequently, while individual species that consistently perform a key role whenever present globally are rare, our results suggest that most species perform a key role in at least a few reefs. 
Indeed, ~70% of all species contributed disproportionally to a function in at least one reef.
Despite high species richness on coral reefs, researchers often report the existence functionally-dominant “key species” [e.g. @Bellwood2006]. 
We demonstrate that while dominance is indeed prevalent, the identity of the dominant species may be less predictable, suggesting that maintaining high levels of species richness across coral reefs is essential to sustain ecosystem functioning at the global scale [@Topor2019].

As many species locally play a dominant role for one or more functions, the vulnerability of these functionally-dominant species heavily affects the vulnerability of the functions on the community scale. 
Species may be more vulnerable to fishing (e.g. large herbivores, predators) or to climate change through coral cover decline (e.g. butterfly fishes) [@Graham2011]. By combining species-level vulnerability scores to fishing and climate change [@Graham2011] with the contributions of each species to the community-scale functions, we estimated the vulnerability of all five functions to both stressors. 
We demonstrate that vulnerability to fishing is relatively the highest for piscivory, followed by P excretion, while climate change vulnerability is the highest for P excretion, N excretion, and production (Fig. 4). 
Double jeopardy tends to be the highest for P excretion and piscivory. 
Surprisingly, since fishing is known to heavily affect large herbivores such as parrot fishes [@Bellwood2011], herbivory comes out as the least vulnerable of all five functions. 
This is a result of the high variety of unique ecosystem roles within the large pool of herbivores. 
Small herbivores are abundant and contribute a lot to herbivory, while they are not as vulnerable. 
Large herbivores may be rarer, particularly in reefs surrounding regions with a high human density, yet they can play crucial roles with low redundancy (e.g. bioerosoion @Bellwood2011).
Thus, our results reflect the overall vulnerability of functions such as herbivory, but in order to conserve specialized functional roles, it is necessary to consider the high vulnerability of specific species that play unique roles [@Bellwood2011]. 
Given the local nature of species’ functional contributions, ecosystem functioning can be supported by local actions, taking into account the dominance and vulnerability, to avert the functional extinction of species [@Mccauley2015], while maintaining a high diversity on a global scale.

Our results quantitatively exemplify why it is challenging to manage Earth’s most diverse and vulnerable marine ecosystems. 
Enhancing ecosystem functioning is an important objective of contemporary conservation initiatives [@Cinner2020]. 
The functional trade-offs revealed between key ecosystem functions challenges the feasibility of this approach and introduces a conundrum for coral reef conservation. 
Traditionally, marine reserves have been the main conservation strategy for coral reefs [@Graham2020].
While they provide benefits for biomass and herbivory, they provide little insurance for diversity in human-dominated regions [@Cinner2020] and do not protect reefs from the pervasive effects of climate change [@Graham2020]. 
Measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions. 
Thus, it is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front]. 
Furthermore, individual reefs require local management plans to protect species with key roles for each function, community structure components, and the presence of multiple stakeholders may require the prioritization of various functions based on local needs and stressors. 
Finally, as most species play an important role on a local scale, local management efforts should go hand in hand with global action to conserve the high diversity of species in addition to tackling climate change, the primary threat to coral reefs [@Bruno2019].

Our results contribute to an ongoing discussion surrounding the question: what constitutes a healthy reef? 
We posit that the answer cannot be rooted in a purely economic mindset that aims for the maximization of all functions. 
Over the past decades, there has been an increasing emphasis on multifunctionality, particularly in terrestrial ecology of plant communities.  
Through the quantification of multiple functions, we exemplify that the concept of multifunctionality may not applicable to hyper-diverse coral reefs.  
Acknowledging and quantifying the complexity of coral reef fish functions is a first step towards a better understanding of reef functioning and vulnerability at the global scale. 
Additionally, there is a need to do this for other groups, such as invertebrates and cryptobenthic fishes that undoubtedly have an important contribution to turnover and fluxes of elements given their small body sizes [@Brandl2019science]. 
Therefore, the majority of coral-reef ecosystem functioning remains to be characterized.
In light of the rapid degradation of coral reefs worldwide [@Hughes2017], quantitative monitoring of multiple functions alongside biomass is critical to create viable conservation targets and uphold coral reef functioning, as well as associated services to humanity, beyond the 21st century [@Brandl2019front].

@@

![Figure 4: Vulnerability of functions based on the cumulative vulnerability of the contribution of species to the five functions per community. Filled bars show the proportion of communities that belong to a certain category of vulnerability. Categories of vulnerability to fishing and climate change are defined using quantiles. Very low: <20%; Low: 20%-40%; Medium: 40%-60%; High: 60%-80%; Very high: > 80%. High vulnerability to both is assigned in the case where both vulnerability to fishing and vulnerability to climate change are either high or very high. Similarly, low vulnerability to both is assigned in the case where both vulnerability to fishing and vulnerability to climate change are either low or very low.](../output/plots/figure_4_vuln_com.png)

>>>>

# Drivers and vulnerability of global coral reef fish functions

Nina M. D. Schiettekatte, Simon J. Brandl, Jordan M. Casey, Nicholas A. J. Graham, Diego R. Barneche, Deron E. Burkepile, Jacob E. Allgeier, Jesús E. Arias-Gonzaléz, Graham J. Edgar, Carlos E. L. Ferreira, Sergio R. Floeter, Alan M. Friedlander, Alison L. Green, Michel Kulbicki, Yves Letourneur, Osmar J. Luiz, Alexandre Mercière, Fabien Morat, Katrina S. Munsterman, Enrico L. Rezende, F. A. Rodríguez‐Zaragoza, Rick D. Stuart-Smith, Laurent Vigliola, Sébastien Villéger, Valeriano Parravicini

**The high diversity and productivity of coral reefs, despite their location in nutrient-poor waters, provides invaluable protection, nutrition, and services to millions of people [@Teh2013]. Protecting reef function (cycling of elements) under the escalating pressures of fishing [@Graham2017; @Ceccarelli2016: @Stuart-Smith2018] and climate change [@Hughes2017] represents the cornerstone of reef conservation. Here, we quantify five key functions—nitrogen and phosphorus cycling, biomass production, herbivory, and piscivory—mediated by reef fishes across the world’s tropical oceans. We demonstrate that critical trade-offs exist among functions (e.g. biomass production and phosphorous cycling), and that no reef fish community worldwide can sustain high values for all five functions simultaneously. Further, functions tend to be dominated by few species, where on average ~12% of species provide more than 50% of a function. However, the identity of these species varies across reefs: ~70% of all species contribute disproportionally to functioning in a given community. Finally, we highlight unbalanced vulnerability across functions, suggesting that certain functions are disproportionally affected by anthropogenic stressors, destabilizing the integrity of coral reef functioning. Our findings introduce a conundrum for coral reef conservation, i.e.we cannot maximize all functions, and stress the need for quantitative monitoring of multiple functions to make informed management decisions.**

The flow of elements through biological communities fuel all life on Earth [@Welti2017]. There is now strong consensus on the importance of preserving these fluxes, often termed ecosystem functions [@Welti2017]. 
For millennia, resources have been managed with an economic mindset to maximize desirable functions such as plant or animal biomass production [@Weisser2017]. 
However, in agro-ecological settings efforts to maximize one function have been shown to negatively impacting another (e.g. timber vs. erosion control), shedding light on the existence of trade-offs between functions [@Tallis2008; @Zavaleta2010]. 
Therefore, sustaining multiple functions, likely requires both high species richness and a variety of species assemblages across the landscape [@Zavaleta2010].
To make informed management decisions, it is critical to quantify multiple functions and their potential trade-offs [@Rodriguez2006]. 
However, beyond well-studied ecosystems such as grasslands, quantifying multiple ecosystem functions simultaneously is challenging, and therefore trade-offs between functions, their drivers, and vulnerability are often poorly understood, impeding informed management [@Brandl2019front].

Coral reefs are among the most diverse and productive ecosystems on earth, and they provide indispensable ecosystem services to humans [@Teh2013]. 
As coral reefs thrive in nutrient-poor waters, efficient cycling of elements is key to their characteristic productivity [@Brandl2019science]. 
Importantly, the integrity of coral reefs is threatened by a plethora of anthropogenic stressors, such as intensive exploitation and climate change [@Hughes2017]. 
Over the past decade, severe declines in coral reef habitat quality and fish biomass, and community structure shifts have brought coral reef functioning and services to the forefront of scientific discourse [@Williams2019; @Bellwood2019; @Brandl2019front].
Yet, our capacity to quantitatively evaluate, monitor, and compare reef functioning primarily relies on static proxies of functions, such as relative live coral cover, fish standing biomass, or functional richness based on qualitative species traits [e.g. @Mora2011plos; @Cinner2020; @Mouillot2014]. 
Conversely, we know comparatively little about elemental fluxes and their drivers (but see [@Allgeier2016]), which currently hinders effective management [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the fluxes of elements that underpin reef fish functioning. 
Using species-specific data on basic organismal processes and phylogenetic extrapolation techniques, we parametrize individual-level bioenergetic models to estimate five key ecosystem functions: nitrogen (N) excretion, phosphorus (P) excretion, biomass production, herbivory (daily consumption of primary producers, expressed in mass of carbon), and piscivory (daily consumption of fishes, expressed in mass of carbon). 
We apply these models to 9,118 reef fish communities across 585 sites worldwide to: 
(1) quantify reef fish functions and their trade-offs, 
(2) extract the community- and species-level predictors of these functions, and 
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

Quantifying five key ecosystem functions performed by fishes across the world's reefs uncovers the functional variation that exists beyond biomass, the most commonly employed indicator of coral reef functioning [@MacNeil2015, @Mora2011plos; @Cinner2020; @Brandl2019front].
Even though we observed a strong relationship between fish standing biomass and all five functions, there  is remarkable variability in reef fish functions beyond biomass (Fig. 1), and for fish communities with similar biomass, functions tend to vary by over two orders of magnitude (Extended Data Fig. 1a). 
Thus, monitoring biomass alone can obscure differences in community-level functioning. 
Further, there are strong trade-offs among the different functions when taking out the effect of biomass (Fig.1, Extended Data Fig. 1b). 
For example, high herbivory rates or N excretion negatively correlate with rates of P excretion.
Strikingly, less than 3% of all communities had an above average biomass-corrected value for all five functions simultaneously, and not a single community reached maximum values (upper 10%) for all functions. 
As a consequence, for a fixed amount of standing biomass, no reef appears to sustain high values for all of the five functions analyzed here. While many reefs may stand out as hotspots for one function, none maximize functioning as a holistic concept (Fig. 1).

Disentangling predictors of reef fish functioning revealed important effects of both community structure and species-specific influences.  
First, we explored the relationship of each function with ecological predictors known to affect elemental fluxes [@Schiettekatte2020] (i.e. body size, trophic level, species richness, biomass, temperature, and age structure; Fig. 2). 
We found that correlations between functions are mediated by contrasting aspects of community structure (Fig. 2). 
For example, P excretion is higher in communities with a high proportion of large-bodied, mature fishes that occupy high trophic levels, thus creating positive relationships between P excretion and piscivory (Extended Fig. 1b) (See also ref. @Allgeier2014). 
In contrast, biomass production is highest in communities dominated by small and/or immature fishes at lower trophic levels, creating a strong trade-off between biomass production and P excretion. 
Metabolic theory predicts that small individuals have a higher mass-specific metabolic rate, and as a consequence, small individuals have elevated consumption rates and disproportionally contribute to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob].
On the other hand, fishes in early life stages that are highly dependent on phosphorus to grow, and fishes with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions to phosphorus excretion. 
These results emphasize the importance of fish community structure for ecosystem-wide functioning [@Schramski2015; @Allgeier2016].

Alongside features of community structure, specific dominant and/or high-performing species strongly affect ecosystem functioning.   
Due to species-specific characteristics, certain species or entire families tend to contribute more or less to a particular function, relative to their contribution to standing biomass (Fig. 3a).
Further, abundant species can profoundly affect rates of functioning at the community level [@Ruttenberg2019; @Topor2019]. 
Therefore, we assessed the role of species by quantifying the degree to which they disproportionally induce (i.e. dominate) different functions for each community. 
We demonstrate that for most communities, functions are consistently dominated by a few species (Fig. 3b). This means for example that on average ~12% of the species inside a community contribute to more than 50% of a function.
However, the identity of these species varies dramatically (Fig. 3c), despite evidence for species assuming key functional roles across widely-dispersed locations [@Bellwood2003].
Thus, while there are almost no globally important species, most species are locally important.
Indeed, ~70% of all species contributed disproportionally to a specific function in at least one reef fish community.
Despite high species richness on coral reefs, researchers often report the existence functionally-dominant “key species” [e.g. @Bellwood2006]. 
We demonstrate that while functional dominance is indeed prevalent, the identity of the dominant species is unpredictable, suggesting that maintaining high levels of species richness across coral reefs is essential to sustain ecosystem functioning at the global scale [@Topor2019].

The critical importance of both reef fish community structure and species-specific contributions shines new light on the vulnerability of coral reef functioning in the Anthropocene. 
Exploitation and anthropogenic climate change have caused severe changes in reef fish biomass and community structure [@Mora2011plos; @Hughes2017; @Williams2019], and our findings suggest strong effects for ecosystem functioning on reefs. 
For example, intensive fishing and associated reductions in biomass of large fishes truncates the size, age, and trophic structure of fish communities [@Graham2017]. 
These effects can enhance biomass-corrected N excretion and production [e.g. @Morais2020glob], while negatively impacting P excretion, herbivory, and piscivory (Fig. 2, Extended Data Fig. 2). 
On the other hand, climate change related declines in coral cover are associated with a shift toward herbivores, which may help avoid algal domination [@Graham2006; @Ceccarelli2016]. 
Yet, herbivores have a minor contribution to phosphorus excretion [@Allgeier2014; @Schiettekatte2020], so a shift to herbivore dominance in fish communities and the subsequent decline of community-level phosphorus excretion may change the balance of nutrient cycling on reefs. 
Higher N:P ratios may in fact favor algal growth [@Burkepile2013], and promote symbiont dominance within the coral holobiont [@Allgeier2020].
This is particularly problematic when considered in conjunction with climate change, since phosphorus starvation in corals reduces their temperature threshold for bleaching [@Ezzat2016]. 
Thus, considering multiple functions paints a more nuanced and holistic picture of how human-induced shifts in community structure affect coral reef functioning.

Similarly, the species-specific vulnerability of functionally-dominant species heavily affects functioning. 
By combining species-level vulnerability scores to fishing and climate change [@Graham2011] with the contributions of each species to each function, we demonstrate that vulnerability to fishing is highest for piscivory, followed by P excretion, while climate change vulnerability is the highest for P excretion, N excretion, and production (Fig. 4), creating a potential double jeopardy of vulnerability for P excretion in particular. 
Surprisingly, although fishing pressure can heavily affect large herbivores such as parrot fishes [@Bellwood2011], herbivory is the least vulnerable function. 
This may be due to the high variability in unique ecosystem roles within the large pool of herbivorous fish species. 
While small herbivores are often abundant and not particularly vulnerable to fishing, larger herbivorous species are often highly targeted and are prone to become functionally extinct in regions with high human density [@Bellwood2012]. 
While herbivores of all body sizes and functional groups are united in our assessment, their realized contributions to herbivory are known to be strongly complementary. 
Thus, our results reflect the overall vulnerability of functions, yet this doesn't exclude potential risk of functional extinction of unique roles by a small group of species, nested in a broad function (e.g. @Bellwood2011).

Conserving biomass, diversity, and ecosystem functioning are important objectives of contemporary conservation initiatives [@Cinner2020]. 
Even though increased fish biomass enhances function, the strong trade-offs we reveal between key ecosystem functions introduces a conundrum for coral reef conservation, where actions to enhance one function may negatively affect another. 
For example, marine reserves, which are the main conservation approach for coral reefs [@Graham2020], may provide benefits for herbivory, but they often offer little insurance for diversity in human-dominated regions [@Cinner2020] and do not protect reefs from the pervasive effects of climate change [@Graham2020]. 
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions. 
Therefore, it is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front], which may leed to informed decisions to prioritize certain functions based on local needs and stressors.  
Finally, while there is a general consensus on the role of diversity in enhancing functioning [@Zavaleta2010; @Hooper2005], we highlight the overarching importance of species identity and dominance at the local scale.
While globally, conserving a high diversity of fishes is indeed beneficial to functioning, on the local scale, species richness only has minor effects on functions.
Rather, community structure and the identity of dominant species tend to play an important role, which emphasizes the importance of monitoring common species to conserve ecosystem functioning [@Gaston2008]. 
Still, as no species provides high contributions for all functions simultaneously, high species diversity enhances performance across multiple functions [@Duffy2009; @Zavaleta2010].

Overall, our results demonstrate that managing coral reefs for ecosystem functioning is, in its simplest form, unlikely to be fruitful. 
While enhancing fish biomass is important, the variability of reef fish communities in the cycling of critical elements is striking. 
Conserving these elemental fluxes in the 21st century will require a more nuanced approach to coral reef management.

@@

![Figure 4: Vulnerability of functions based on the cumulative vulnerability of the contribution of species to the five functions per community. Filled bars show the proportion of communities that belong to a certain category of vulnerability. Categories of vulnerability to fishing and climate change are defined using quantiles. Low: <20%; High: >80%. High vulnerability to both is assigned in the case where both vulnerability to fishing and vulnerability to climate change are high. Similarly, low vulnerability to both is assigned in the case where both vulnerability to fishing and vulnerability to climate change are low.](../output/plots/figure_4_vuln_com.png)

### 2020-09-24 text/methods.Rmd
commit before deep clean

## 1. Underwater visual census database
\noindent We used a published global database of reef fish abundance and sizes collected along belt transects [@Barneche2019]. This database encompasses 9118 transects across 585 sites (98 localities) in the Central Indo-Pacific, Central Pacific, Eastern Pacific, Western Indian, Eastern Atlantic, Western Atlantic. The database only includes sites with a hard reef bottom at the outer reef slope. Transects were carried out at a constant depth, parallel to the reef crest. We selected the species inside families for which we have body stoichiometric data, that were at least 7cm to minimize the bias related to the indentification of small individuals, and finally we discarded rare species, for which less than 20 individuals were ever recorded across all transects. The dataset then included 1110 species that belong to 25 families (Acanthuridae, Balistidae, Bothidae, Chaetodontidae, Cirrhitidae, Fistulariidae, Haemulidae, Holocentridae, Kyphosidae, Labridae, Lethrinidae, Lutjanidae, Monacanthidae, Mugilidae, Mullidae, Ostraciidae, Pempheridae, Pomacanthidae, Pomacentridae, Sciaenidae, Scorpaenidae, Serranidae, Siganidae, Tetraodontidae, Zanclidae).

@@

\noindent Finally, we quantified baseline values for each function and assessed the occurrence of communities that reach baseline values for each function. Based on the biomass reference for a healthy reef of ~100g/m2 (@MacNeil2015), we predicted the baseline for each function for the full range of sst's. Then, for all communities, we quantified whether or not the community had higher values for all functions compared to their baselines. Using this logical variable, we fitted a Bernoulli model with a logit-link to predict the probability of having all five functions above their threshold in relation to biomass and sea surface temperature, similar to the model specified above.

## 4. Effect of community structure on ecosystem functions 
\noindent To investigate the effect of the community structure accounting for the effect of standing biomass and sea surface temperature, we quantified a set of variables that characterize the community. These variables discribe the size, age, and trophic distribution of the community, as these may all affect functions [@Schiettekatte2020]. Specifically, we took the 2.5%, 50% and 97.5% quantiles of the total length, immaturity, and trophic level of all individuals per transect. The total length is based on the visual estimation by divers. The immaturity is quantified using the following formula: 
<!-- \begin{equation} -->
	$$\textrm{immaturity}_\textrm{i} = \kappa (\textrm{l}_{\infty} - \textrm{l}_\textrm{i}),$$
<!-- 	\label{imm} -->
<!-- \end{equation} -->
where $\kappa$ is the species-specific growth rate parameter and $l_{\infty}$ is the species-specific asymptotic adult length, and $l_i$ is the total length of individual i. Essentially, this is the derivative of the Von Bertalanffy growth model for a certain length, and the higher this value is, the younger the individual. Finally, the trophic level was extracted from fishbase [@Froese2018]. Additionally, we quantified the transect-level species richness. 
For each log-transformed function we then fitted a Bayesian mixed-effect model with all 12 above-mentioned variables:

@@

## 6. Species vulnerability to fishing and clilate change
For each species, we quantified two measures of vulnerability: vulnerability to climate change and vulnerability to fishing pressure [@Graham2011]. For species' vulnerability to climate change, we solely focus on their vulnerability to the loss of live corals. Vulnerability to climate change is related to diet specialization, habitat specialization, recruitment specialization for live coral and body size [@Graham2011]. Graham et al. (2011) [@Graham2011] developed a score for climate change vulnerability for 134 species. We used these scores to fit a Bayesian mixed effect predictive model that relates the vulnerability with the log-transformed maximum size of fish (extracted from Fishbase @Froese2018), the dependence on coral for food (3 categories: not dependent, facultative corallivore, and obligate corallivore), and dependence on coral for habitat (2 categories: dependent vs. not dependent). We also included a random effect for family. This model had a Bayesian R2 of 0.97. We then used this model to extrapolate the vulnerability measure to all 1110 species in our dataset. For species' vulnerability to fishing, we extracted the index from Cheung et al. (2005) @Cheung2005.

>>>>

## 1. Underwater visual census database
\noindent We used a published global database of reef fish abundance and sizes collected along belt transects [@Barneche2019]. This database encompasses 9118 transects across 585 sites (98 localities) in the Central Indo-Pacific, Central Pacific, Eastern Pacific, Western Indian, Eastern Atlantic, Western Atlantic. The database only includes sites with a hard reef bottom at the outer reef slope. Transects were carried out at a constant depth, parallel to the reef crest. We selected the species inside families for which we have body stoichiometric data, that were at least 7cm to minimize the bias related to the identification of small individuals, and finally we discarded rare species, for which less than 20 individuals were ever recorded across all transects. The dataset then included 1110 species that belong to 25 families (Acanthuridae, Balistidae, Bothidae, Chaetodontidae, Cirrhitidae, Fistulariidae, Haemulidae, Holocentridae, Kyphosidae, Labridae, Lethrinidae, Lutjanidae, Monacanthidae, Mugilidae, Mullidae, Ostraciidae, Pempheridae, Pomacanthidae, Pomacentridae, Sciaenidae, Scorpaenidae, Serranidae, Siganidae, Tetraodontidae, Zanclidae).

@@

## 4. Effect of community structure on ecosystem functions 
\noindent To investigate the effect of the community structure accounting for the effect of standing biomass and sea surface temperature, we quantified a set of variables that characterize the community. These variables describe the size, age, and trophic distribution of the community, as these may all affect functions [@Schiettekatte2020]. Specifically, we took the 2.5%, 50% and 97.5% quantiles of the total length, immaturity, and trophic level of all individuals per transect. The total length is based on the visual estimation by divers. The immaturity is quantified using the following formula: 
<!-- \begin{equation} -->
	$$\textrm{immaturity}_\textrm{i} = \kappa (\textrm{l}_{\infty} - \textrm{l}_\textrm{i}),$$
<!-- 	\label{imm} -->
<!-- \end{equation} -->
where $\kappa$ is the species-specific growth rate parameter and $l_{\infty}$ is the species-specific asymptotic adult length, and $l_i$ is the total length of individual i. Essentially, this is the derivative of the Von Bertalanffy growth model for a certain length, and the higher this value is, the younger the individual. Finally, the trophic level was extracted from fishbase [@Froese2018]. Additionally, we quantified the transect-level species richness. 
For each log-transformed function we then fitted a Bayesian mixed-effect model with all 12 above-mentioned variables:

@@

## 6. Vulnerability to fishing and climate change
For each species, we quantified two measures of vulnerability: vulnerability to climate change and vulnerability to fishing pressure [@Graham2011]. For species' vulnerability to climate change, we solely focus on their vulnerability to the loss of live corals. Vulnerability to climate change is related to diet specialization, habitat specialization, recruitment specialization for live coral and body size [@Graham2011]. Graham et al. (2011) [@Graham2011] developed a score for climate change vulnerability for 134 species. We used these scores to fit a Bayesian mixed effect predictive model that relates the vulnerability with the log-transformed maximum size of fish (extracted from Fishbase @Froese2018), the dependence on coral for food (3 categories: not dependent, facultative corallivore, and obligate corallivore), and dependence on coral for habitat (2 categories: dependent vs. not dependent). We also included a random effect for family. This model had a Bayesian R2 of 0.97. We then used this model to extrapolate the vulnerability measure to all 1110 species in our dataset. For species' vulnerability to fishing, we extracted the index from Cheung et al. (2005) @Cheung2005.
Then, we calculated vulnerability scores per function on the community level by averaging the species-level scores weighted by the contributions to function of species. Finally, we calculated proportions of communities with either high (upper quartile) or low (lower quartile) vulnerability for each function. We also report the proportion of communities that have high or low vulnerability scores for both climate change and fishing.
